---
title: "EDS241: Assignment 3"
author: "Scou Leonard"
date: '`r format(Sys.time(), "%m/%d/%Y")`'
output: 
  pdf_document:
    toc: false
    number_sections: yes
header-includes:
  - \setlength{\parindent}{1em}
  - \usepackage{float}
--- 

``` {r setup, echo = FALSE, message = FALSE, warning = FALSE}
# set default chunk options
knitr::opts_chunk$set(fig.width = 4, fig.height = 3, 
                      echo = TRUE, message = FALSE, warning = FALSE)


# load packages
packages=c("stargazer", 
           "here", 
           "tidyr", 
           "dplyr",
           "stringr", 
           "janitor", 
           "cowplot", 
           "ggplot2", 
           "tinytex", 
           "datasets", 
           "tibble",
           "here",
           "tidyverse",
           "estimatr",
           "huxtable")

for (i in packages) {
  if (require(i,character.only=TRUE)==FALSE) {
    install.packages(i,repos='http://cran.us.r-project.org')
  }
  else {
    require(i,character.only=TRUE)
  }
}

#devtools::install_github('rstudio/rmarkdown')
options(scipen=999) # not scientific notation
```

\noindent This exercise asks you to implement some of the techniques presented in Lectures 6-7. The goal is to estimate the causal effect of maternal smoking during pregnancy on infant birth weight using the treatment ignorability assumptions. The data are taken from the National Natality Detail Files, and the extract “SMOKING_EDS241.csv”' is a random sample of all births in Pennsylvania during 1989-1991. Each observation is a mother-infant pair. The key variables are:

**The outcome and treatment variables are:**

- `birthwgt` = birth weight if infant in grams
- `tobacco` = indicator for maternal smoking 

**The control variables are:**

- `mage`: mother's age
- `meduc`: mother's education
- `mblack`: = 1 if mother is Black
- `alcohol`: =1 if consumed alcohol during pregnancy
- `first`: =1 if first child
- `diabete`: = 1 if mother is diabetic
- `anemia`: =1 if mother anemic 

# Load and Clean Data 

```{r}
#read in the data
smoking_df <- read.csv(here("data", "SMOKING_EDS241.csv"))
```

\newpage 

# Homework Questions

## Part A

\noindent What is the unadjusted mean difference in birth weight of infants with smoking and non- smoking mothers? Under what assumption does this correspond to the average treatment effect of maternal smoking during pregnancy on infant birth weight? Provide some simple empirical evidence for or against this hypothesis.

```{r}
mod_a1 <- lm_robust(birthwgt ~ tobacco, data = smoking_df)

huxreg(mod_a1)
```

\noindent The unadjusted mean difference in birth weight of infants with smoking and non- smoking mothers is -244.5 grams. 

\noindent This tells us about the effect of smoking on infant birth weight assuming that mothers who smoke and mothers who do are statistically different. 

```{r}
mod_a2 <- lm_robust(meduc ~ tobacco, data = smoking_df)

huxtable::huxreg(mod_a2)
```

\noindent Evidence against this hypothesis is shown in the code chunk above; when we regress tobacco usage on education, the results tell us that there is a significantly significant difference between the education levels of mothers who use tobacco during pregancy and those who do not. From this results, we know that mother's education is correlated with tobacco usage, but it is not included in our prediction of birthweights of infants of mothers who use tobacco above. This means there might be omitted variables bias, and the there may be additional variables interacting with the condition of tobacco use which are excluded. 

\newpage

## Part B

\noindent Assume that maternal smoking is randomly assigned conditional on the observable covariates listed above. Estimate the effect of maternal smoking on birth weight using a linear regression. Report the estimated coefficient on tobacco and its standard error.

\noindent The code chunk below estimates the effect of of maternal smoking on birth weight using a linear regression.

```{r}
#linear regression with birthweight conditional on all of the variables 
mod_B <- lm_robust(birthwgt ~ ., 
                   data = smoking_df)

#standard error
mod_B[[2]][[2]]
```

\noindent The average treatment effect of maternal smoking on birth weight when all other covariants are held equal is -228.07 grams (on average, infants of mothers who use tobacco weigh 236.46 grams less than the infants of mothers who do not), with a standard error of `r round(mod_b[[2]][[2]], digits = 2)`.

\newpage

## Part C

\noindent Use the exact matching estimator to estimate the effect of maternal smoking on birth weight. For simplicity, consider the following covariates in your matching estimator: create a 0-1 indicator for mother's age (=1 if mage>=34), and a 0-1 indicator for mother's education (1 if meduc>=16), mother's race (mblack), and alcohol consumption indicator (alcohol). These 4 covariates will create 2*2*2*2 = 16 cells. Report the estimated average treatment effect of smoking on birthweight using the exact matching estimator and its linear regression analogue (Lecture 6, slides 12-14).

\noindent The code chunk below creates 0-1 indicators for mother's education and mother's age.

```{r}
#create indicators for mage and meduc 
smoking_df <- smoking_df %>% 
  mutate(mage_indicator = case_when(mage >= 34 ~ 1,
                                    mage < 34 ~ 0)) %>% 
  mutate(meduc_indicator = case_when(meduc >= 16 ~ 1,
                                     meduc < 16 ~ 0)) %>% 
  mutate(g = paste0(mage_indicator, 
                    meduc_indicator, 
                    mblack,
                    alcohol)) %>% 
  mutate(g = as.factor(g))
```

\noindent The code chunk below generates the linear regression analogue to estimate the effect of smoking on birth weight. 

```{r}
#LINEAR REG ANALOGUE 
mod_C <- lm_robust(birthwgt ~  tobacco +
                     mage_indicator + 
                     meduc_indicator + 
                     mblack + 
                     alcohol + 
                     mage_indicator:meduc_indicator +
                     mage_indicator:mblack +
                     mage_indicator:alcohol +
                     meduc_indicator:mblack +
                     meduc_indicator:alcohol +
                     mblack:alcohol +
                     mage_indicator:meduc_indicator:mblack +
                     mage_indicator:meduc_indicator:alcohol +
                     meduc_indicator:mblack:alcohol +
                     mage_indicator:meduc_indicator:mblack:alcohol, 
                   data = smoking_df)
```

```{r}
#generate table of coefficients for mod_C
huxtable::huxreg(mod_C)
```

\noindent The linear regression analogue tells us that the the infants of mothers who smoke during pregnancy weigh 226.25 grams less than the infants of mothers who do not smoke during pregnancy.

```{r}
#EXACT MATCHING ESTIMATOR
TIA_table <- smoking_df %>%
  group_by(g,tobacco) %>% 
  summarise(n_obs = n(),
            birthwgt_mean= mean(birthwgt, 
                                na.rm = TRUE )) %>% 
  gather(variables, values, 
         n_obs:birthwgt_mean) %>% #Reshape data
  mutate(variables = paste0(variables,"_",tobacco, sep="")) %>%
  pivot_wider(id_cols = g, names_from = variables,values_from = values) %>% 
  ungroup() %>%  #Ungroup from X values
  mutate(birthwgt_diff = birthwgt_mean_1 - birthwgt_mean_0, #calculate Y_diff
         w_ATE = (n_obs_0+n_obs_1)/(sum(n_obs_0)+sum(n_obs_1)),
         w_ATT = n_obs_1/sum(n_obs_1)) %>% #calculate weights
  mutate_if(is.numeric, round, 2) #Round data


stargazer(TIA_table, type= "text", summary = FALSE, digits = 2)

# MULTIVARIATE MATCHING ESTIMATES OF ATE AND ATT
ATE=sum((TIA_table$w_ATE)*(TIA_table$birthwgt_diff))
ATE
```

\noindent The average treatment effect generated using the exact matching estimator is -224.26. 

\newpage 

## Part D

\noindent Estimate the propensity score for maternal smoking using a logit estimator and based on the following specification: mother’s age, mother’s age squared, mother’s education, and indicators for mother’s race, and alcohol consumption.

```{r eval =FALSE}
smoking_df <- smoking_df %>% 
  mutate(mage_2 = mage * mage)

# BASIC PROPENSITY SCORE --- THIS IS A TOY MODEL
# ESTIMATE PROPENSITY SCORE MODEL AND PREDICT (EPS)
ps_model <- glm(tobacco ~ mage + mage_2 + meduc + mblack + alcohol, family = binomial(), data = smoking_df)

EPS <- predict(ps_model, type = "response")
PS_WGT <- (smoking_df$tobacco/EPS) + ((1-smoking_df$tobacco)/(1-EPS))
```

\newpage

## Part E

\noindent Use the propensity score weighted regression (WLS) to estimate the effect of maternal smoking on birth weight (Lecture 7, slide 12).

```{r}
wls1 <- lm_robust(birthwgt ~ tobacco, data = smoking_df, weights = PS_WGT)

huxtable::huxreg(wls1)
```

\noindent Using the propensity score weighted regression, the effect of maternal smoking in birth weight is that infants born to mothers who use tobacco during pregnancy weigh 225.48 grams less than infants born to mothers who do not use tobacco. 
